### 2016-04-20 ALMA/Cycle4/RXJ.tex
with Phil's comments

While the ISM in local ULIRGs has been studied in great detail,
forming a rich inventory of molecular transitions that serves as the template for 
studying high-z galaxies 
\citep[e.g.][]{Rangwala15a}, a wide knowledge gap persists
between $z$=0 and the epoch when most stars are formed in the universe ($z$$\sim$2). 
Hence, understanding galaxy populations at an epoch when
the build-up of stellar mass 
is steeply rising ($z$$\sim$0.3-1) is critical and we here aim to bridge this gap by testifying
correlations and properties found locally out to high redshift by observing the 
dynamical structure and properties of the ISM of
the quadruply lensed galaxy RXJ1131-1231 and its dust-obscured 
companion at $z_{\rm CO}$$\sim$0.65 (\Fig{HST}).

@@

While \comol emission traces the total molecular distribution and dynamics,
high-dipole moment molecules such as HCN and \hcop are expected to trace the 
%properties of the 
denser, actively star-forming gas. Indeed, a tight correlation between \ahcn and \LIR 
(proxy for SFR) has been found in nearby galaxies and local giant molecular 
clouds (GMC; \citealt[hereafter GS04]{Gao04a}; \citealt{Wu05}), 
suggesting HCN is a faithful tracer of the star-forming dense gas.
% and that the origin can be attributed to the fundamental units of star formation. 
However, \dhcn observations of (U)LIRGs have revealed a wide range of
excitation conditions 
%for their dense gas phase 
that may render the ground state transition of 
HCN and \hcop poor proxies of the dense gas mass \citep[hereafter P07]{Papadopoulos07a}.
In this light, higher-$J$ transitions (e.g. $J$=4\rarr3) 
% are necessary to scrutinize the reliability of using the ground state 
have been suggested to be better probes
% dense gas mass tracers since the mid-$J$ transitions
since they trace the much denser gas (n$\gtrsim$10$^5$-10$^7$ cm$^{-3}$) that is 
thought to be the immediate fuel for SF in turbulent GMCs \citep{Shirley03a, KM05}. 
This is supported by the linear correlations found in 
\LIR-\Lp[\dhcn] and \LIR-\Lp[\dhcop] \citep[]{Zhang14a}.
Since the 
ground state lines are redshifted to frequencies beyond the spectral coverage of ALMA
at $z$$>$0.06, it is also necessary to establish diagnostics using these mid-$J$ lines to 
study distant galaxies.
Additionally, due to the difference in abundances and excitation conditions of HCN and \hcop
in star-forming versus AGN regions, the line 
% ratios \ahcn/HCO$^+$($J$=1\rarr0) \citep{Kohno05a, Imanishi10, Izumi13a} and
ratio \dhcn/\dhcop have been 
proposed as
%  diagnostic tools 
a diagnostic tool to reveal deeply buried AGNs at the cores of ULIRGs \citep{Imanishi14a, GB14a, Viti14a, Izumi16a, Imanishi16a}.
% separate emission originating from AGN and SB
% Recent studies have
% extended this diagnostic to using the J=4\rarr3 transitions 
%\citep{Imanishi14a, GB14a, Viti14a}, 
% where an elevated line ratio is seen in AGNs, suggesting
% it can be used to reveal deeply buried AGNs at the cores of ULIRGs \citep{Izumi16a, Imanishi16a}.

Prior to ALMA, studies of dense gas were largely limited to the local universe ($z$$\lesssim$0.1) 
with only five detections
%IR-luminous lensed galaxies detected 
at $z$$\gtrsim$0.3
(e.g. \citealt{Riechers07a, Riechers10a, Wagg05a, Gao07a}).
Moreover, none of these 
spatially resolve the emission, rendering it difficult 
to draw more detailed conclusions on the dense gas properties of galaxies at high z. 
Even with ALMA, such studies will remain challenging, but by combining 
the magnification provided by gravitational lensing with the exceptional spatial resolution and sensitivity of ALMA, 
studies
of dense molecular gas in distant galaxies are now possible, as proposed here.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section*{Proposed Observations and Immediate Objectives}
% \vspace{-0.25em}
\noindent {\bf Proposed Observations and Science Goals}
We here propose to map (1): \eco at 0.15" resolution ($\sim$500pc at $z$$\sim$0.7 in the source plane) and 
(2): \dhcn and \dhcop emission and the underlying continuum at 0.7" resolution (2.5 kpc in the source plane).
The continuum traces the dust emission, 
providing better constraints on the the gas-to-dust ratio, dust temperature(s), dust mass,
and its spatial extent (and thus the surface density of the SFR).
%($\Sigma_{\rm SFR}$).
These quantities are key to investigate how the ISM % conditions 
vary as galaxies evolve across cosmic time.
In conjunction with the large set of ancillary data from
rest-frame UV to radio wavelength and our recent observations of \bco and \rot[CO]{3}{2}, our
proposed observations are designed to investigate 
% carry a detailed investigation of 
the gas-rich merger in aspects listed as follows.

\underline{\bf Dynamics and kinematics:}
Our recently obtained \bco data shows an asymmetric double-horned line profile (\Fig{combine}a). Given
the 1$^{\rm st}$ moment map and the observed velocity gradient
across the source plane in our model (\Fig{combine}d \& \ref{fig:combine}f), this is indicative of
a kinematically ordered %(reminiscent of a rotating disk) 
galaxy, but its emission has been lensed differentially. 
Limited by the spatial resolution of this data, it is not possible to infer the true 
kinematics due to beam smearing.
Furthermore, the unusually high velocity dispersion 
$\gtrsim$400\kms at the central region (\Fig{combine}e) hints at 
perturbations from the AGN and/or
internal turbulence
% motion 
due to interactions with the companion
and/or instability due to the huge gas reservoir.
Thus higher-resolution \eco imaging, as proposed here, is needed
to distinguish 
between a merger-driven and a turbulent clumpy disk starburst/

@@

\underline{\bf Line ratios and the AGN/SB diagnostic:}
Aided by lensing, we will be able obtain kinematical information on spatial 
scales smaller than the beam, as seen in our \bco data (\Fig{combine}c, d), thereby
% and thus recover structural details and kinematics in finer detail than otherwise possible
enabling us to probe the physical properties of
the inner molecular disk near the AGN. 
% At the proposed resolution, we will be able to distinguish the \dhcn and \dhcop
% emission originating from the AGN host galaxy (and within) and from its companion.
Recent studies find that the \dhcn/\dhcop line ratio is expected to be enhanced 
in the circumnuclear disk (CND) near an AGN and falls off with distance from the CND
\citep{GB14a,Imanishi16a}. At the proposed resolution, we will 
resolve the line emission originating near the AGN and that from the spatially extended 
SB regions ($>$kpc scales) in the host galaxy, and by measuring spatial variations in 
this line ratio, we will assess its utility 
as an AGN/SB diagnostic (\Fig{AGNSB}) at high redshift.
Since the HCN vibrational line (v$_2$=1;$J$=4\rarr3) falls within the targeted frequency range, 
we will use it to independently constrain the radiation properties of the AGN environment, 
given that this line is excited by infrared pumping and that the radiation
field associated with the AGN is the cause of the elevated \dhcn/\dhcop
\citep{Sakamoto10a,Imanishi13a}.
The proposed spatial resolution is also necessary for
a detailed lens modeling of each emission line, enabling measurements of reliable line 
ratios without significant uncertainties due to differential lensing.
Such measurements are vital to reveal how galaxy interactions drive gas into 
inner disks to initiate SB/AGN activities and how the ISM differs from 
normal star-forming galaxies.

@@

\noindent Our proposed observations therefore provide an exceptional opportunity to investigate
the physical properties and dynamical structures of different gas phases in the ISM of mergers 
and distant galaxies with exquisite details. 
This will also demonstrate the capabilities of ALMA to % mapping 
utilize these much fainter
high-dipole moment molecules as routine tracers to 
study star-formation at earlier cosmic epochs.

@@

\begin{figure*}[!tbph]
\centering
\vspace{-0.559em}
\hspace{-2.75em}
\floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=0.445\textwidth}}]{figure}[\FBwidth]
{
\includegraphics[trim=7 6 5 5, clip, width=0.575\textwidth]{Fig/combine}}
{
\hspace{-1.0em}
\caption{ \fontsize{10pt}{12pt}\selectfont {\textbf{ New \bco} data.}
{\em (a):}
A double-horned line profile from the AGN host galaxy, which appears
 asymmetric due to differential lensing.
{\em (b):}
Spectra taken at three locations along the strongest velocity gradient, demonstrating
differential lensing of the kinematic components of the gas-rich ``disk".
{\em (c):}
Observed spatial variations across the channels, as shown by the red (red-shifted), 
green (line center), and blue (blue-shifted) contours.
{\em (d, e):}
The observed velocity gradient and velocity dispersion are suggestive of a kinematically ordered disk at 
the current resolution limit. 
The spectrally resolved lensed emission allows us to probe dynamical structures on smaller spatial 
scales than otherwise possible.
{\em (f):}
Channel maps of the CO emission (red) 
overlaid on our best-fit lens models (grayscale). 
The foreground lensing galaxy is represented by a black dot.
The reconstructed source morphology (magenta ellipses) is also suggestive of a ``disk".
Given the presence of a companion galaxy within 2.4 kpc and beam smearing effects, high-resolution 
imaging is necessary to unambiguously determine the structure that gives rise to the 
observed velocity gradient and dispersion, as proposed here.
This will allow us to investigate the 
mechanisms responsible for the starburst in the AGN host galaxy
% (e.g. GMC-like in a rotating disk or fragmentation of a
% dynamically unstable gas-rich disk) 
and its ISM conditions as it interacts with the companion. 
We therefore aim to spatially resolve the gas dynamics, kinematics and ISM conditions in a merger,
providing observational constraints 
on the star-formation processes
at an epoch where the SFR density is steeply 
rising across cosmic times.
\label{fig:combine}}
}
\vspace{-0.905em}
\end{figure*}
%

>>>>

While the ISM in local ULIRGs has been studied in great detail,
forming a rich inventory of molecular transitions that serves as the template for 
studying high-z galaxies 
\citep[e.g.][]{Rangwala15a}, a wide knowledge gap persists
between $z$=0 and the epoch when most stars are formed in the universe ($z$$\sim$2). 
Hence, understanding galaxy populations at an epoch when
the build-up of stellar mass 
is steeply rising ($z$$\sim$0.3-1) is critical and we here aim to bridge this gap by testing
correlations and properties found locally out to high redshift by observing the 
dynamical structure and properties of the ISM of
the quadruply lensed galaxy RXJ1131-1231 (henceforth RX1131) and its dust-obscured 
companion at $z_{\rm CO}$$\sim$0.65 (\Fig{HST}).

@@

While \comol emission traces the total molecular distribution and dynamics,
high-dipole moment molecules such as HCN and \hcop are expected to trace the 
%properties of the 
denser, actively star-forming gas. Indeed, a tight correlation between \ahcn and \LIR 
(proxy for SFR) has been found in nearby galaxies and local giant molecular 
clouds (GMC; \citealt[hereafter GS04]{Gao04a}; \citealt{Wu05}), 
suggesting HCN is a faithful tracer of the star-forming dense gas.
% and that the origin can be attributed to the fundamental units of star formation. 
However, \dhcn observations of (U)LIRGs have revealed a wide range of
excitation conditions 
%for their dense gas phase 
that may render the ground state transition of 
HCN and \hcop poor proxies of the dense gas mass \citep[hereafter P07]{Papadopoulos07a}.
In this light, higher-$J$ transitions (e.g. $J$=4\rarr3) 
% are necessary to scrutinize the reliability of using the ground state 
have been suggested to be better probes
% dense gas mass tracers since the mid-$J$ transitions
since they trace the much denser gas (n$\gtrsim$10$^5$-10$^7$ cm$^{-3}$) that is 
thought to be the immediate fuel for SF in turbulent GMCs \citep{Shirley03a, KM05}. 
This is supported by the linear correlations found in 
\LIR-\Lp[\dhcn] and \LIR-\Lp[\dhcop] \citep[]{Zhang14a}.
Since the 
ground state lines are redshifted to frequencies beyond the spectral coverage of ALMA
at $z$$>$0.06, it is also necessary to establish diagnostics using these mid-$J$ lines to 
study distant galaxies.
Additionally, due to the difference in abundances and excitation conditions of HCN and \hcop
in star-forming versus AGN regions, the line 
% ratios \ahcn/HCO$^+$($J$=1\rarr0) \citep{Kohno05a, Imanishi10, Izumi13a} and
ratio \dhcn/\dhcop has been 
proposed as
%  diagnostic tools 
a diagnostic tool to reveal deeply buried AGNs at the cores of ULIRGs \citep{Imanishi14a, GB14a, Viti14a, Izumi16a, Imanishi16a}.
% separate emission originating from AGN and SB
% Recent studies have
% extended this diagnostic to using the J=4\rarr3 transitions 
%\citep{Imanishi14a, GB14a, Viti14a}, 
% where an elevated line ratio is seen in AGNs, suggesting
% it can be used to reveal deeply buried AGNs at the cores of ULIRGs \citep{Izumi16a, Imanishi16a}.

Prior to ALMA, studies using dense gas tracers were largely limited 
to the local universe ($z$$\lesssim$0.1) with only five detections
%IR-luminous lensed galaxies detected 
at $z$$\gtrsim$0.3
(e.g. \citealt{Riechers07a, Riechers10a, Wagg05a, Gao07a}).
Moreover, none of these 
spatially resolve the emission, rendering it difficult 
to draw more detailed conclusions on the dense gas properties of galaxies at high z. 
Even with ALMA, such studies will remain challenging, but by combining 
the magnification provided by gravitational lensing with the exceptional spatial resolution and sensitivity of ALMA, 
studies
of dense molecular gas in distant galaxies are now possible, as proposed here.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section*{Proposed Observations and Immediate Objectives}
% \vspace{-0.25em}
\noindent {\bf Proposed Observations and Science Goals}
We here propose to map (1): \eco at 0.15" resolution ($\sim$500pc at $z$$\sim$0.7 in the source plane) and 
(2): \dhcn and \dhcop emission and the underlying continuum at 0.7" resolution (2.5 kpc in the source plane).
The continuum traces the dust emission, 
providing better constraints on the the gas-to-dust ratio, dust temperature(s), dust mass,
and its spatial extent (and thus the surface density of the SFR).
%($\Sigma_{\rm SFR}$).
These quantities are key to investigate how the ISM % conditions 
varies as galaxies evolve across cosmic time.
In conjunction with the large set of ancillary data from
rest-frame UV to radio wavelength and our recent observations of \bco and \rot[CO]{3}{2}, our
proposed observations are designed to investigate 
% carry a detailed investigation of 
the gas-rich merger in aspects listed as follows.

\underline{\bf Dynamics and kinematics:}
Our recently obtained \bco data shows an asymmetric double-horned line profile (\Fig{combine}a). Given
the 1$^{\rm st}$ moment map and the observed velocity gradient
across the source plane in our model (\Fig{combine}d, f), this is indicative of
a kinematically ordered %(reminiscent of a rotating disk) 
galaxy, but its emission has been lensed differentially. 
Limited by the spatial resolution of this data, it is not possible to infer the true 
kinematics due to beam smearing.
Furthermore, the unusually high velocity dispersion 
$\gtrsim$400\kms in the central region (\Fig{combine}e) hints at 
perturbations from the AGN and/or
internal turbulence
% motion 
due to interactions with the companion
and/or instability due to the huge gas reservoir.
Thus higher-resolution \eco imaging, as proposed here, is needed
to distinguish 
between a merger-driven and a turbulent clumpy disk starburst.

@@

\underline{\bf Line ratios and the AGN/SB diagnostic:}
Aided by lensing, we will be able obtain kinematical information on spatial 
scales smaller than the beam, as seen in our \bco data (\Fig{combine}c, d), thereby
% and thus recover structural details and kinematics in finer detail than otherwise possible
enabling us to probe the physical properties of
the inner molecular disk near the AGN. 
% At the proposed resolution, we will be able to distinguish the \dhcn and \dhcop
% emission originating from the AGN host galaxy (and within) and from its companion.
Recent studies find that the \dhcn/\dhcop line ratio is enhanced 
in the circumnuclear disk (CND) near an AGN and falls off with distance from the CND
\citep{GB14a,Imanishi16a}. At the proposed resolution, we will 
resolve the line emission originating near the AGN and that from the spatially extended 
SB regions ($>$kpc scales) in the host galaxy, and by measuring spatial variations in 
this line ratio, we will assess its utility 
as an AGN/SB diagnostic (\Fig{AGNSB}) at high redshift.
Since the HCN vibrational line (v$_2$=1;$J$=4\rarr3) falls within the targeted frequency range, 
we will use it to independently constrain the radiation properties of the AGN environment, 
given that this line is excited by infrared pumping and that the radiation
field associated with the AGN is the cause of the elevated \dhcn/\dhcop
\citep{Sakamoto10a,Imanishi13a}.
The proposed spatial resolution is also necessary for
a detailed lens modeling of each emission line, enabling measurements of reliable line 
ratios without significant uncertainties due to differential lensing.
Such measurements are vital to reveal how galaxy interactions drive gas into 
inner disks to initiate SB/AGN activities and how the ISM differs from 
normal star-forming galaxies.

@@

\medskip
\noindent In summary, our proposed observations therefore provide an exceptional opportunity to investigate
the physical properties and dynamical structures of different gas phases in the ISM of a distant merger
% and distant galaxies 
at exquisite detail. 
This will also demonstrate the capabilities of ALMA to % mapping 
utilize these much fainter
high-dipole moment molecules as routine tracers to 
study star-formation at earlier cosmic epochs.

@@

\begin{figure*}[!tbph]
\centering
\vspace{-0.559em}
\hspace{-2.75em}
\floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=0.445\textwidth}}]{figure}[\FBwidth]
{
\includegraphics[trim=37 35 247 5, clip, width=0.575\textwidth]{Fig/combine.pdf}}
{
\hspace{-1.0em}
\caption{ \fontsize{10pt}{12pt}\selectfont {\textbf{Recent \bco data from NOEMA.}}
{\em (a):}
A double-horned line profile from the AGN host galaxy, which appears
 asymmetric due to differential lensing.
{\em (b):}
Spectra taken at three locations along the strongest velocity gradient, demonstrating
differential lensing of the kinematic components of the gas-rich ``disk".
{\em (c):}
Observed spatial variations across the channels, as shown by the red (red-shifted), 
green (line center), and blue (blue-shifted) contours.
{\em (d, e):}
The observed velocity gradient and velocity dispersion are suggestive of a kinematically ordered disk at 
the current resolution limit. 
The spectrally resolved lensed emission allows us to probe dynamical structures on smaller spatial 
scales than otherwise possible.
{\em (f):}
Channel maps of the CO emission (red) 
overlaid on our best-fit lens models (grayscale). 
The foreground lensing galaxy is represented by a black dot.
The reconstructed source morphology (magenta ellipses) is also suggestive of a ``disk".
Given the presence of a companion galaxy within 2.4 kpc and beam smearing effects, high-resolution 
imaging is necessary to unambiguously determine the structure that gives rise to the 
observed velocity gradient and dispersion, as proposed here.
This will allow us to investigate the 
mechanisms responsible for the starburst in the AGN host galaxy
% (e.g. GMC-like in a rotating disk or fragmentation of a
% dynamically unstable gas-rich disk) 
and its ISM conditions as it interacts with the companion. 
We therefore aim to spatially resolve the gas dynamics, kinematics and ISM conditions in a merger,
providing observational constraints 
on the star-formation processes
at an epoch where the SFR density is steeply 
rising across cosmic times.
\label{fig:combine}}
}
\vspace{-0.905em}
\end{figure*}
%
